---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


### Set the environment before data processing
```{r, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
local <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_ALL", "English") 
```


## Loading and processing data from the zipped file
```{r, echo=TRUE}
con <- unz("activity.zip", "activity.csv")
df <- read.csv(con)
df$date <- ymd(df$date)
```


## What is mean total number of steps taken per day?
1. Histogram of the total number of steps taken each day
```{r, echo=TRUE}
df1 <- df %>% group_by(date) %>% summarise(Total_steps=sum(steps, na.rm = T))
ggplot(df1, aes(x=date, y= Total_steps))+ geom_col()
s_mean= mean(df1$Total_steps)
s_median= median(df1$Total_steps)

```
2. Report results  
The mean total number of steps taken per day is `r s_mean` steps   
The median total number of steps taken per day is `r s_median` steps  

## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?




### Set the environment back
```{r, results='hide'}
Sys.setlocale(locale=local)
```