---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
editor_options: 
  chunk_output_type: console
---


## Set the environment before data processing
```{r, message=FALSE, warning=FALSE, results='hide'}
library(tidyverse)
library(lattice)
local <- Sys.getlocale("LC_TIME")
Sys.setlocale("LC_ALL", "English") 
```


## Loading and processing data from the zipped file
```{r, echo=TRUE}
con <- unz("activity.zip", "activity.csv")
ori_df <- read.csv(con)
df <- ori_df
```


## What is mean total number of steps taken per day?
**Histogram of the total number of steps taken each day**
```{r, echo=TRUE}
df$date <- ymd(df$date)
p <- df %>% group_by(date) %>% summarise(Total_steps=sum(steps, na.rm = T))
hist(p$Total_steps, breaks = 30, xlab="Total number of steps taken in a day", 
     main = "Histogram of the total number of steps taken in a day")

s_mean= mean(p$Total_steps)
s_median= median(p$Total_steps)
```
**Results**  
The mean total number of steps taken per day is `r s_mean` steps   
The median total number of steps taken per day is `r s_median` steps  

## What is the average daily activity pattern?
**In 5-minute interval**  
```{r}
df$interval <- as.numeric(df$interval)
p <- df %>% group_by(interval) %>% summarise(Average_steps= mean(steps, na.rm =TRUE))

#ggplot(p, aes(x=interval, y=Average_steps))+ geom_line()
xyplot(Average_steps~interval, data = p, type="l")
```

**Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?**  
```{r}
x <- p[which(p$Average_steps== max(p$Average_steps)),]
x
```
**Result:**  
The `r x[, 1]` interval contains the maximum number `r x[,2]` of steps.

## Imputing missing values
**Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)**  
```{r}
total_NA <- sum(is.na(df))

```
**Results:**  
The total number of NAs in the dataset is `r total_NA`

**Use the mean steps for that day to fill in all of the missing values in the dataset.**  

```{r}
df2 <- df %>% group_by(date) %>% summarise(day_mean=mean(steps, na.rm = TRUE))
df2$date <- ymd(df2$date)
df2 <- left_join(df, df2, by="date")

for (i in 1:nrow(df2)){
        if (is.na(df2$steps[i]))  {
                if (is.na(df2$day_mean[i])){
                        df2$steps[i] <- 0
                }else {
                        df2$steps[i] <- df2$day_mean[i]
                }
        }else{
                next
        }
}

new_df <- ori_df
new_df$steps <- df2$steps
# this new_df is the same as the original dataset except the Nas are filled
new_df$date <- ymd(new_df$date)
p <- new_df %>% group_by(date) %>% summarise(Total_steps=sum(steps, na.rm = T))
hist(p$Total_steps, breaks = 30, xlab="Total number of steps taken in a day", 
     main = "Histogram of the total number of steps taken in a day")
s_mean= mean(p$Total_steps)
s_median= median(p$Total_steps)
s_mean
s_median
```
**Results:**  
1. The new dataframe I make is called new_df. It's the same as the original dataset except the NAs in column steps are filled  
2. The calculated mean for the new dataset is `r s_mean`  
The calculated median for the new dataset is `r s_median`    
3. The way I used to fill the missing value by either the mean steps of the day or 0 if the value of the whole day is missing, basically produces the same result as the original dataset.  

## Are there differences in activity patterns between weekdays and weekends?

```{r}
#Make the weekend/weekday indicator (factor)
df$steps <- new_df$steps
df <- df %>% mutate (week_day= ifelse(wday(date)== 7| wday(date)==1, "weekend", "weekday"))
df$week_day <- factor(df$week_day, levels=c("weekday", "weekend"))


#Comparing weekends and weekdays on the average of steps with 5-minute interval
p <- df %>% group_by(interval, week_day) %>% summarise(Average_steps= mean(steps))

#if use ggplot
# ggplot(p, aes(x=interval, y=Average_steps))+ geom_line()+ facet_grid(week_day~.)

xyplot(Average_steps~interval|week_day, data=p, type='l', layout=c(1, 2),
       scales=list(x=list(at=c(0500, 1000, 1500, 2000), labels=c(500, 1000, 1500, 2000))))
```
  
**Results:**  
There is greater variation of activities (steps) in weekdays than on weekends, with
the highest average total steps in a 5-minute interval happened in the morning of weekdays around interval 800-900.  


## Set the environment back
```{r, results='hide'}
Sys.setlocale(locale=local)
```